The problem and promise of scale in multilayer networks

### 2020-02-02 paper/manuscript.Rmd
cite todos, fills and little fixes to ml description

Our study is motivated by behavioural, landscape, and spatial ecology to build a
conceptual and analytical framework for scale-dependent multilayer networks and
we apply this framework to a case study of caribou (*Rangifer tarandus*) in
Newfoundland, Canada (Peignier et al. 2019). We use this case study to address
two principal questions associated with scale-dependent multilayer networks.
First, to what degree does the social scale of relationships, spatial scale of
landscapes, and temporal scale of analysis influence our ability to interpret
complex social systems? Second, can variation in social, spatial, and temporal
improve predictions associated with biological differences in seasonal resource
availability and social association? We conclude with perspectives for further
development and application of scale in multilayer networks with respect to
landscape ecology (Schneider 2009), individual fitness and phenotypes (Webber &
Vander Wal 2018), and movement ecology and collective movement (Jolles et al.
2019).

@@

### Caribou socioecology
Here, we use caribou as a case study to develop scale-dependent multilayer
networks based on social (visual and auditory perception), spatial (habitat
type), and temporal (seasonal). Caribou are gregarious ungulates with
fission-fusion dynamics (Lesmerises et al. 2018) that display temporal
[@Peignier_2019] and spatial (Webber et al. unpublished data) variation in
social networks. In winter, caribou dig holes in the snow, termed craters, to
access forage (Bergerud 1974). As a result, caribou tend to occupy and re-use
craters once they are established and there is considerably less access to
forage than when the landscape is snow-free. Importantly, craters exist on the
landscape at multiple scales: the crater scale (a single crater), the feeding
area scale (multiple craters in close proximity), and the winter range scale
(all craters within an individual’s range) [@Mayor_2009]. The distribution of
forage for caribou is therefore seasonally heterogeneous, with greater access in
snow-free seasons compared to winter as well as spatially heterogeneous in
winter because the distribution of craters on the landscape varies meaning that
access to vegetation in winter is highly variable for caribou. We therefore
partitioned data into two discrete seasons:  summer (3 August - 20 September)
and winter (1 January – 17 February) based on caribou socioecology
[@Peignier_2019].

### Caribou location data
We used GPS location data collected from Fogo Island caribou in 2017 and 2018.
Adult female caribou were immobilized and fitted with global positioning system
(GPS) collars (Lotek Wireless Inc., Newmarket, ON, Canada, GPS4400M collars,
1,250 g) as described by Schaefer and Mahoney (2013). Collars were programmed to
collect location fixes every 2 hours. Prior to analyses, we subset GPS fixes
into a discrete removed all erroneous and outlier GPS fixes following Bjørneraas
et al. (2010). We did not collar all female caribou in the herd; however, we
assumed that our sample of collared animals was random.

@@

#### Network types and construction (defining layers for the multilayer network)

<!-- should use something like "using syntax from Kivela 2014..." or 
Following [@Kivela_2014]:
other info to maybe include: the multilayer network was not node-aligned since
not all individuals were on all layers
the temporal aspect is ordinal coupling and landcover aspect is categorical coupling

Individuals were subset to ensure they had data collected in both seasons

-->

@@

Within each habitat-specific monolayer network, we assumed individuals were
associating if simultaneous GPS fixes (i.e., recorded within 5 minutes of each
other) were within a given distance of one another. Typically for ungulates and
other gregarious mammals, the ‘chain rule’ is applied for group assignment. For
GPS data, the chain rule is applied if discrete spatiotemporal GPS fixes
buffered by a given distance formed a contiguous buffer for two or more
individuals, even if some individuals within the buffer were not within the
given distance of one another [@Robitaille_2019]. Group assignment based
on the chain rule has previously been applied to caribou at a distance of 50m
[@Peignier_2019; @Lesmerises_2018]. In all networks, nodes
represented individuals, intralayer edges represented associations between
individuals in a given habitat type, and interlayer edges represented
connections between the same individuals across contexts.

We weighted edges of social networks by the strength of association between
caribou using the simple ratio index (SRI, Cairns and Schwager 1987):

@@

where x is the number of fixes where individuals A and B were within a given
distance of each other, $y_{A}$ is the number of fixes from individual A
when individual B did not have a simultaneous fix, $y_{B}$ is the number
of fixes from individual B when individual A did not have a simultaneous fix,
and $y_{AB}$ is the number of simultaneous fixes from individuals A and B
that were separated by more than the given distance (Farine & Whitehead 2015).
Social groups were designated if two or more individuals occurred within a given
distance of one another at any given time point.

>>>>

Our study is motivated by behavioural, landscape, and spatial ecology to build a
conceptual and analytical framework for scale-dependent multilayer networks and
we apply this framework to a case study of caribou (*Rangifer tarandus*) in
Newfoundland, Canada [@Peignier_2019]. We use this case study to address
two principal questions associated with scale-dependent multilayer networks.
First, to what degree does the social scale of relationships, spatial scale of
landscapes, and temporal scale of analysis influence our ability to interpret
complex social systems? Second, can variation in social, spatial, and temporal
improve predictions associated with biological differences in seasonal resource
availability and social association? We conclude with perspectives for further
development and application of scale in multilayer networks with respect to
landscape ecology (Schneider 2009)<!-- TODO-->, individual fitness and phenotypes (Webber &
Vander Wal 2018), and movement ecology and collective movement (Jolles et al.
2019).

@@

### Caribou socioecology
Here, we use caribou as a case study to develop scale-dependent multilayer
networks based on social (visual and auditory perception), spatial (habitat
type), and temporal (seasonal). Caribou are gregarious ungulates with
fission-fusion dynamics [@Lesmerises_2018] that display temporal
[@Peignier_2019] and spatial (Webber et al. unpublished data) variation in
social networks. In winter, caribou dig holes in the snow, termed craters, to
access forage (Bergerud 1974)<!-- TODO-->. As a result, caribou tend to occupy and re-use craters once they are established and there is considerably less access to
forage than when the landscape is snow-free. Importantly, craters exist on the
landscape at multiple scales: the crater scale (a single crater), the feeding
area scale (multiple craters in close proximity), and the winter range scale
(all craters within an individual’s range) [@Mayor_2009]. The distribution of
forage for caribou is therefore seasonally heterogeneous, with greater access in
snow-free seasons compared to winter as well as spatially heterogeneous in
winter because the distribution of craters on the landscape varies meaning that
access to vegetation in winter is highly variable for caribou. We therefore
partitioned data into two discrete seasons:  summer (3 August - 20 September)
and winter (1 January – 17 February) based on caribou socioecology
[@Peignier_2019].

### Caribou location data
We used GPS location data collected from Fogo Island caribou in 2017 and 2018.
Adult female caribou were immobilized and fitted with global positioning system
(GPS) collars (Lotek Wireless Inc., Newmarket, ON, Canada, GPS4400M collars,
1,250 g) as described by Schaefer and Mahoney (2013). Collars were programmed to
collect location fixes every 2 hours. Prior to analyses, we subset GPS fixes
into a discrete removed all erroneous and outlier GPS fixes following Bjørneraas
et al. (2010)<!-- TODO-->. We did not collar all female caribou in the herd; however, we
assumed that our sample of collared animals was random.

@@

#### Network types and construction (defining layers for the multilayer network)
<!-- should use something like "using syntax from Kivela 2014..." or 
Following [@Kivela_2014]:
other info to maybe include: the multilayer network was not node-aligned since
not all individuals were on all layers
the temporal aspect is ordinal coupling and landcover aspect is categorical coupling

Individuals were subset to ensure they had data collected in both seasons
-->

@@

Within each layer, we assumed individuals were
associating if simultaneous GPS fixes (i.e., recorded within 5 minutes of each
other) were within a given distance of one another. Typically for ungulates and
other gregarious mammals, the ‘chain rule’ is applied for group assignment. For
GPS data, the chain rule is applied if discrete spatiotemporal GPS fixes
buffered by a given distance formed a contiguous buffer for two or more
individuals, even if some individuals within the buffer were not within the
given distance of one another [@Robitaille_2019]. Group assignment based
on the chain rule has previously been applied to caribou at a distance of 50m
[@Peignier_2019; @Lesmerises_2018]. In all networks, nodes
represented individuals, intralayer edges represented associations between
individuals in a given habitat type and season, and interlayer edges represented
connections between the same individuals across contexts.

We weighted edges of social networks by the strength of association between
caribou using the simple ratio index (SRI, Cairns and Schwager 1987): <!-- TODO-->

@@

where x is the number of fixes where individuals A and B were within a given
distance of each other, $y_{A}$ is the number of fixes from individual A
when individual B did not have a simultaneous fix, $y_{B}$ is the number
of fixes from individual B when individual A did not have a simultaneous fix,
and $y_{AB}$ is the number of simultaneous fixes from individuals A and B
that were separated by more than the given distance [@Farine_2015].
Social groups were designated if two or more individuals occurred within a given
distance of one another at any given time point.
